%
% \documentclass[acmsmall,screen, review, nonacm]{acmart}
\documentclass[10pt]{article} % For LaTeX2e
\usepackage{amsmath}
\usepackage{amssymb} % For \mathbb and other symbols
\usepackage{amsfonts} % For additional math fonts

% \usepackage{tmlr}
\usepackage[preprint]{tmlr}

\title{\Tool{}: LLM Generation with Grammar Augmentation}

% For Theorem
\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\usepackage{thm-restate}
% \declaretheorem[name=Definition]{definition}
% \declaretheorem[name=Lemma]{lemma}
% \declaretheorem[name=Theorem]{theorem}

% \declaretheorem{theorem}

\usepackage{hyperref}
\usepackage{pifont}
\usepackage{threeparttable} % for tablenotes

% Algo related
\usepackage{algorithm}

% \usepackage{algorithmic}
\usepackage[noend]{algpseudocode} % for some features in algorithm
\usepackage{mathtools} % for '\smashoperator' macro
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{caption, subcaption}
\usepackage{multirow}
 \usepackage{setspace}
 \usepackage[normalem]{ulem}
 \usepackage{enumitem}
 \usepackage{wrapfig}
 \usepackage{bm}
 \usepackage{xspace}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{csquotes}
%\usepackage{amssymb}
% \usepackage{float}
\usepackage{listings}
\usepackage{xcolor} % For custom colors
\usepackage{soul}
% \usepackage{lstlinebgrd}

% To switch between ICLR and Arxiv versions
\usepackage{ifthen}
\newboolean{anonymous}
\setboolean{anonymous}{false}

\input{defs}
\newboolean{showcomments}
\setboolean{showcomments}{true}
\input{commenting}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{} % Clears all page headers
\renewcommand{\headrulewidth}{0pt}

\begin{document}
%
% \title{\Tool{}: LLM Generation with Grammar Augmentation}
%

\ifthenelse{\boolean{anonymous}}
{
}
{
\author{\name Shubham Ugare \\
       \addr University of Illinois Urbana-Champaign, USA
       \AND
       \name Tarun Suresh \\
       \addr University of Illinois Urbana-Champaign, USA
       \AND
       \name Hangoo Kang \\
       \addr University of Illinois Urbana-Champaign, USA
       \AND
       \name Sasa Misailovic \\
       \addr University of Illinois Urbana-Champaign, USA
       \AND
       \name Gagandeep Singh \\
       \addr University of Illinois Urbana-Champaign and VMware Research, USA
}
}

\maketitle              

\begin{abstract}
LLMs are widely used in complex AI applications.
These applications underscore the need for LLM outputs to adhere to a specific format, for their integration with other components in the systems.
Typically the format rules -- e.g., data serialization formats such as JSON, YAML, or Code in Programming Language -- are expressed as context-free grammar (CFG).
Due to the hallucinations and unreliability of LLMs, instructing LLMs to adhere to specified syntax becomes an increasingly important challenge.


We present \Tool{}, a novel framework for efficient and general syntactical decoding with LLMs, to address this challenge. 
% precision
\Tool{} ensures soundness and completeness with respect to the CFG of a formal language, effectively retaining valid tokens while filtering out invalid ones.
% effciency
\Tool{} uses an offline-constructed, efficient lookup table, the \textit{DFA mask store}, created from the DFA (Deterministic Finite Automaton) of the language’s grammar for efficient generation.
% framework
\Tool{} seamlessly integrates with any language defined by CFG, as evidenced by experiments focusing on generating JSON, SQL, Python, and Go outputs. 
Our experiments evaluating the effectiveness of \Tool{} for JSON generation demonstrate that \Tool{} eliminates all syntax errors and significantly outperforms state-of-the-art baselines.
Furthermore, our results underscore how \Tool{} significantly reduces \average{} of syntax errors in generated Python and Go code, showcasing its substantial impact on enhancing syntactical precision in LLM generation.

\ifthenelse{\boolean{anonymous}}{

}
{
% \vspace{0.1in}
Our code is available at {\color{blue}\url{https://github.com/uiuc-focal-lab/syncode}}
}



% \keywords{Large\,Language\,Model \and\! Syntax\,Error \and{}Context-free\,Grammar}
\end{abstract}


\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Background}
\label{sec:background}
\input{background}

\section{Overview}
\label{sec:overview}
\input{overview}

\section{\mbox{Syntactically Correct Generation}}
\label{sec:technical}
\input{technical/technical}

\section{Experimental Methodology}
\input{eval}

\section{Related Work}
\label{sec:related}
\input{related}

\section{Conclusion}
\input{conclusion}

% \input{limitations}
%
% ---- Bibliography ----
%
\bibliographystyle{ACM-Reference-Format}
\bibliography{ref}

\include{Appendix}
\end{document}
