\documentclass{article}
\usepackage{iclr2025_conference,times}
\iclrfinalcopy
% \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
% \pagestyle{plain} % removes running headers

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}
 \usepackage{wrapfig}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{enumitem}
\usepackage{multirow}
% \usepackage{censor}

% To switch between ICLR and Arxiv versions
\usepackage{ifthen}
\newboolean{iclr}
\setboolean{iclr}{true}


% For code blocks
\usepackage{tcolorbox}
\tcbuselibrary{listings}
% \tcbuselibrary{listingsutf8}

% \tcbuselibrary{minted}
\tcbuselibrary{theorems,listings,skins,breakable}

% Define colors for listings
\definecolor{keywordcolor}{rgb}{0.67,0.13,1.00}
\definecolor{commentcolor}{rgb}{0.15,0.5,0.15}
\definecolor{stringcolor}{rgb}{0.75,0.10,0.10}

% Configure listings for colored syntax
%Define custom EBNF language for syntax highlighting
% Configure listings for colored syntax
\lstset{
  keywordstyle=\color{keywordcolor}\bfseries,
  commentstyle=\color{commentcolor}\itshape,
  stringstyle=\color{stringcolor},
  basicstyle=\ttfamily\small,
  breaklines=true,
  keywords={start, paragraph, sentence, word, punctuation},
}

%Define custom EBNF language for syntax highlighting
\lstdefinestyle{EBNF}{
    keywords={},
    keywordstyle=\color{blue}\bfseries,
    basicstyle=\ttfamily\scriptsize,
    comment=[l]{\%},
    commentstyle=\color{orange},
    moredelim=[is][\color{blue}\bfseries]{**}{**}
}

\lstdefinestyle{pythonstyle}{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\bfseries\color{blue},        % Keywords like def, class, etc.
    commentstyle=\color{gray}\ttfamily,        % Comment style
    stringstyle=\color{red},                   % String style
    showstringspaces=false,
    tabsize=4,                                 % Set tab size (4 spaces is typical for Python)
    frame=single,
    breaklines=true,
    emph={generate_sql_with_itergen},                               % Emphasize specific Python keywords like 'self'
    emphstyle=\color{orange},                  % Color for emphasized words like 'self'
    morekeywords={as, None, True, False},      % Add Python-specific keywords that might be missing
    otherkeywords={:, !=, =, +, *, -, /, //, +=, ==, >=, <=},  % Add operator highlighting
    literate={\-}{}{0\discretionary{-}{}{-}},  % Break long words by hyphenation
}


\setlist{leftmargin=5.5mm}

% For Theorem
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{thm-restate}

\usepackage{graphicx}
\usepackage{caption, subcaption}

\declaretheorem[name=Definition]{definition}
\declaretheorem[name=Lemma]{lemma}
\declaretheorem[name=Theorem]{theorem}


\input{defs}
% \input{commenting}
\title{IterGen: Iterative Semantic-aware Structured LLM Generation with Backtracking}


\author{Shubham Ugare,\ \ Rohan Gumaste,\ \ Tarun Suresh,\ \ \textbf{Gagandeep Singh},\ \ Sasa Misailovic  \\
University of Illinois Urbana-Champaign\\
\{\texttt{sugare2}, \texttt{gumaste2}, \texttt{tsuresh3}, \texttt{ggnds}, \texttt{misailo}\}\texttt{@illinois.edu} \\
}


%\includeonly{}
% \includeonly{appendix}

\begin{document}


\maketitle


\begin{abstract}

Large Language Models (LLMs) are widely used for tasks such as natural language and code generation. Still, their outputs often suffer from issues like privacy violations and semantically inaccurate code generation. 
Current libraries for LLM generation rely on left-to-right decoding without systematic support for backtracking, limiting the ability to correct or refine outputs mid-generation. 

To address this issue, we introduce \Tool{}, an intuitive framework for iterative, grammar-guided LLM generation that enables users to move both forward and backward within the generated output based on grammar symbols. 
By leveraging a symbol-to-position mapping, \Tool{} ensures efficient and structured generation while allowing for corrections during the process. 
We demonstrate \Tool{}'s effectiveness in two important applications: reducing privacy leakage in LLM outputs, improving the accuracy of LLM-generated SQL and Vega-Lite queries.

\vspace{0.1in}
Our code and additional resources are available at~{\color{blue}\tt\footnotesize\href{https://structuredllm.com}{http://structuredllm.com}}.

% {\color{blue}\url{https://github.com/uiuc-arc/itergen}} and 
% Our code is available at {\color{blue}\url{https://github.com/uiuc-arc/itergen}}

% \ifthenelse{\boolean{iclr}}
% {
% }
% {
% \vspace{0.1in}
% Our code is available at {\color{blue}\url{https://github.com/uiuc-arc/itergen}}
% }

\end{abstract}

\vspace{-.02in}
\section{Introduction}
\label{intro}
\input{intro}

\vspace{-.05in}
\section{Background}
\input{background}

\section{\mbox{Iterative structured generation}}
\vspace{-.05in}
\input{itergen}

% \section{Experimental Methodology}
% \input{methodology}
\vspace{-.15in}

\section{Evaluation}
\input{eval}

\vspace{-.15in}

\section{Related Work}
\label{sec:related}
\input{related}

\vspace{-.05in}
\section{Limitations}
\vspace{-.15in}
\input{limitations}

\vspace{-.1in}
\section{Conclusion}
\vspace{-.1in}
We present \Tool{}, an efficient and general framework that uses the symbols in the BNF grammar for intuitive iteration over the LLM generation of structured outputs. It brings the flexibility of bidirectional iterators from standard programming languages to LLM-based generation. 

By enabling users to enforce syntactic and semantic constraints, \Tool{} significantly advances {the reliability of LLM outputs.} 
%
Our evaluation already demonstrates its effectiveness in improving text-SQL generation on average by 18.5\% over existing state-of-the-art techniques and fully eliminating privacy leaks in LLM-generated text. 
Furthermore, by enforcing semantic constraints, \Tool{} improves the accuracy of LLM-generated Vega-Lite specifications by 17.8\%.
% \sasa{TBD: Add the results for Vega-light}
%
In the future, we anticipate that \Tool{} presents the solid foundation for easily expressing and enforcing various complex semantic properties of structured texts, including code, documents, and natural language, during generation with open-source LLMs. 



\newpage 

\ifthenelse{\boolean{iclr}}
{
\section{Reproducibility Statement} We provide the source code of \Tool{} as part of the supplementary material that can be used to reproduce our results. We also provide additional experimental details and pseudocode of the algorithm in the appendix.

}
{
}


\section*{ACKNOWLEDGMENTS}
We thank the anonymous reviewers for their comments. This research work was supported in part by NSF Grants No. CCF-2238079, CCF-2316233, CNS-2148583, CCF-1846354, CCF-2313028 and the IBM-Illinois Discovery Accelerator Institute.

\bibliographystyle{plainnat}
\bibliography{ref}

\appendix

\newpage

\include{appendixa}

\end{document}
