
    \begin{tcblisting}{
        listing only, 
        halign=left,
        listing engine=listings,
        title=\textbf{\small Prompt for Vega-lite case study},
        colbacktitle=blue!30!white, 
        coltitle=black,
        left=0pt,               % Left inner padding
        right=0mm,              % Right inner padding
        top=0mm,                % Reduce top margin
        bottom=0mm,             % Reduce bottom margin
        boxsep=1mm,
        width=\textwidth,
        listing options={style=EBNF},
        label={fig:overview_grammar},
    }
    You are an expert AI model in data visualization, skilled at converting natural language descriptions into Vega-Lite JSON specifications. Vega-Lite is a high-level JSON-based visualization grammar for creating interactive and multi-view visualizations. Its specifications describe a single or complex composed view, using properties such as mark (visual type) and encoding (mapping data fields to visual properties). Each JSON specification should begin with the following structure. 
    
"$schema": "https://vega.github.io/schema/vega-lite/v3.json",
"data": {
    "url": "datasets/{dataset}.csv"
}
    
Given a natural language request, output a Vega-Lite JSON object that meets the request requirements. Only include the "$schema", "data", "mark", and "encoding" keys in the JSON object.

For example:
Request: "Show a bar chart of the number of houses in each city."
Dataset: houses
Data fields: "City", "Price", "Size"
Vega-Lite JSON Specification:
{
    "$schema": "https://vega.github.io/schema/vega-lite/v3.json",
    "data": {
        "url": "datasets/houses.csv"
    },
    "mark": {"type": "bar"},
    "encoding": {
        "x": {"field": "City", "type": "nominal"},
        "y": {"aggregate": "count", "type": "quantitative", "axis": {"title": "COUNT"}}
	}
}

Each JSON object should accurately reflect the query's intent, using appropriate Vega-Lite encoding, marks, and transformations. Use "datasets/{dataset}.csv" as the data source. Can you convert the given utterance into a VEGA-Lite specification?
Utterance: Scatterplot mpg vs displacement color by origin
Dataset: cars
Data fields: Model, MPG, Cylinders, Displacement, Horsepower, Weight, Acceleration, Year, Origin
Vega-Lite JSON Specification:
\end{tcblisting}
\label{fig:database_schema_query}

